---
title: "Optimization Project: INFO 2020"
author: "Ryan Elmore"
date: "2/17/2019"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(janitor)
library(lpSolve)
knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## Instructions
Solve the following optimization problems to the best of your ability. You must
work individually on your solutions and may only ask me (Ryan) for help. Each
person has a unique problem and your answers will *not* match any other 
student's solutions. If there is evidence that you are submitting the solution
to your friend's problem and not your own, you will receive a zero for the 
project. 

## Background
The Denver Nuggets (away team) were playing at the Los Angeles Lakers (home 
team) on October 25, 2018. I have collected data and built a model to project
how many fantasy points each player is going to score during this game. We 
are going use optimization methods to select a collection of six players 
players in order to maximize the expected points that they will earn during
this game. Each part of the problem is subject to a unique set of 
constraints. The general problem is a variant of a fantasy sports problem. 

Each successive problem must satisfy the following contraints, in addition to
the constraints set out in each part of the project. 

* Your "team" of six players may not exceed more than \$50,000 in total 
salary. 
* You must have at least one player from each team, *i.e.* at least one 
player from the Lakers and at least one from the Nuggets. 

### Problem 1
Find the optimal set of six players in order to maximize expected points 
subject to the salary constraint mentioned above. 

```{r data, include = FALSE}
set.seed(10982)
date_string <- "251018"
df <- read_csv(paste("../data/DKSalaries-",
                        date_string,
                        ".csv",
                        sep = "")) %>%
  clean_names() %>%
  filter(roster_position == "CPT") %>%
  separate(game_info, into = c("away", "rest"), sep = "@") %>%
  mutate(home = if_else(away == team_abbrev, 0, 1),
         salary = salary/1.5) %>%
  separate(position, c("p1", "p2")) %>%
  select(name, team = team_abbrev, position = p1, home_indicator = home,
         salary, avg_points_per_game) 
df <- df[1:20, ] %>%
  mutate(expected_points = round(avg_points_per_game + rnorm(20))) %>%
  select(-avg_points_per_game)
```


```{r, include = FALSE}
obj <- df$expected_points
tm_mat <- t(model.matrix(~ team + 0, df))
player_mat <- t(model.matrix(~ name + 0, df))
con <- rbind(rep(1, nrow(df)),
             df$salary,
             tm_mat,
             # rep(df$home, 2),
             # rep(df$is.starter, 2),
             player_mat)

dir <- c("=", "<=", rep("<=", 2), rep("<=", nrow(df)))
rhs <- c(6, 50000, rep(5, 2), rep(1, nrow(df)))
result <- lp("max", obj, con, dir, rhs, all.bin = TRUE)
opt_team <- df[result$solution == 1, ]
opt_salary <- sum(opt_team$salary)
```

The expected points for original optimal lineup is `r result$objval`. 

We used `r opt_salary` of the 50000 in salary for this lineup. 

The team consists of the following players:
```{r}
opt_team$name
```

### Problem 2

Suppose we want at least five Lakers on our proposed team of six players. What
is the optimial team in terms of maximizing expected points? 

```{r}
obj <- df$expected_points
tm_mat <- t(model.matrix(~ team + 0, df))
player_mat <- t(model.matrix(~ name + 0, df))
con <- rbind(rep(1, nrow(df)),
             df$salary,
             tm_mat,
             df$team == "LAL",
             # rep(df$home, 2),
             # rep(df$is.starter, 2),
             player_mat)

dir <- c("=", "<=", rep("<=", 2), ">=", rep("<=", nrow(df)))
rhs <- c(6, 50000, rep(5, 2), 5, rep(1, nrow(df)))
result <- lp("max", obj, con, dir, rhs, all.bin = TRUE)
opt_team <- df[result$solution == 1, ]
opt_salary <- sum(opt_team$salary)
```

The expected points for original optimal lineup is `r result$objval`. 

We used `r opt_salary` of the 50000 in salary for this lineup. 

The team consists of the following players:
```{r}
opt_team$name
```

### Problem 3

On the other hand, the Nuggets are a better team. Remove the constraint from
part 2 and add a constraint saying that you want at least four players from 
the Nuggets. What is the optimial team in terms of maximizing expected points?

```{r}
obj <- df$expected_points
tm_mat <- t(model.matrix(~ team + 0, df))
player_mat <- t(model.matrix(~ name + 0, df))
con <- rbind(rep(1, nrow(df)),
             df$salary,
             tm_mat,
             df$team == "DEN",
             # rep(df$home, 2),
             # rep(df$is.starter, 2),
             player_mat)

dir <- c("=", "<=", rep("<=", 2), ">=", rep("<=", nrow(df)))
rhs <- c(6, 50000, rep(5, 2), 4, rep(1, nrow(df)))
result <- lp("max", obj, con, dir, rhs, all.bin = TRUE)
opt_team <- df[result$solution == 1, ]
opt_salary <- sum(opt_team$salary)
```

The expected points for original optimal lineup is `r result$objval`. 

We used `r opt_salary` of the 50000 in salary for this lineup. 

The team consists of the following players:
```{r}
opt_team$name
```


### Problem 4

On the other hand, the Nuggets are a better team. Remove the constraint from
part 2 and add a constraint saying that you want at least four players from 
the Nuggets. What is the optimial team in terms of maximizing expected points?

```{r}
obj <- df$expected_points
tm_mat <- t(model.matrix(~ team + 0, df))
player_mat <- t(model.matrix(~ name + 0, df))
con <- rbind(rep(1, nrow(df)),
             df$salary,
             tm_mat,
             df$team == "DEN",
             # rep(df$home, 2),
             # rep(df$is.starter, 2),
             player_mat)

dir <- c("=", "<=", rep("<=", 2), ">=", rep("<=", nrow(df)))
rhs <- c(6, 50000, rep(5, 2), 4, rep(1, nrow(df)))
result <- lp("max", obj, con, dir, rhs, all.bin = TRUE)
opt_team <- df[result$solution == 1, ]
opt_salary <- sum(opt_team$salary)
```

The expected points for original optimal lineup is `r result$objval`. 

We used `r opt_salary` of the 50000 in salary for this lineup. 

The team consists of the following players:
```{r}
opt_team$name
```

